Initial Analysis

### 2019-05-23 Initial_Analysis.Rmd
Ready to run in Beast

```{r setup, eval=T, echo=F, warning=F,message=F, results='hide'}

@@

#### Library ####
packages <- c(
  "readxl", # Read dataframe
  "data.table", # Read dataframe (Fast!)
  "dplyr", # Data manipulation
  "tidyr", # Data manipulation
  "ggplot2", #Nice grpahs and spatial analysis
  # "cowplot",
  # "rgdal",
  # "RColorBrewer",
  # "knitr",
  # "kableExtra",
  # "ggrepel",
  # "gridExtra",
  # "ggmap",
  # "rgeos",
  # "stringr",
  # "stringer",
  "spdep", # for poly2nb
  "sf", #Spatial analysis 
  "sp", #Spatial analysis 
  "purrr",#Spatial analysis
  "rgdal", #Spatial analysis
  "tools"#, #Spatial analysis 
  # "png", # For reading plots in chunk codes
  # "grid" # For reading plots in chunk codes
)

@@

SAU <- length(list.files("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/SAU_Distribution/"))
ENM <- length(list.files("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/ENM/"))
Occ <- length(list.files("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/Occurence/"))

@@

@@

EEZ_Size <- EEZ_CellID %>% 
  group_by(EEZID) %>% 
  summarise(n_grid = n()) %>% 
  left_join(EEZIDs_List,
            by = "EEZID")

@@

  # SAU Distributions
  if(Model == "SAU"){
    File_Name <- paste("SAU_Distribution/DIST_GOOD_SP_")
  }

@@

    Load <- lapply(Models_List, FUN=R.matlab::readMat, na.strings=0)
    sppdist <- as.data.frame(bind_cols(Load))
    sppdist$INDEX <- INDEX
    colnames(sppdist) <- c("SAU","Occ","ENM","INDEX")

@@

# Variables outside function
# head(getSppDist(600004,"SAU")) # Works
# head(getSppDist(600004,"Occurence")) # Works
# head(getSppDist(600004,"ENM")) # Works
# head(getSppDist(600004,"All")) # Works

@@

```{r Trans_Spp_Function, eval=T,echo=T,warning=F,message=F}

@@

# The function
EstTransIndex <- function(SppDist,EEZ_Size,Neighbors){

@@

  Trans_Spp <- SppDist %>%
    # gather("Model","Value",1:3) %>% 
    filter(
      INDEX %in% Index_EEZ$INDEX,
      # Value != 0
      SAU != 0
      ) %>%
    left_join(Neighbors,
              by = "INDEX") %>% # Match the species distribution to the Countries EEZs
    group_by(Name = Country_Territory,
             Neighbor_Territory#,
             # Model
             ) %>% 
    summarise(n_spp = n()) %>% # The number of cells present within each country's EEZ
    left_join(EEZ_Size,
              by = "Name") %>% 
    mutate(Trans_Index = n_spp/n_grid) %>%  # Include one transboundary species
    select(EEZID,everything())  # For the sake of my brain

  Countries_with_Specie <- Trans_Spp %>% 
    filter(Trans_Index == 1) %>% 
    group_by(Neighbor_Territory) %>% 
    summarise()

  # Function Results

@@

### Testing Function ####
# SppDist <- Spp_Dist
# EEZ_Size <- EEZ_Size
# 
# Neighbors <- Neighbor_List

@@

# Routine

@@

#### TESTING PERIOD ####
# I'm substetting central america to do the whole analysis (smaller EEZs = faster computation) once I have figure out things I'll run for all

@@

## Species Test Data

# Species list

# DBEM exploited species list 
exploited_species_list <- fread("/Volumes/DATA/PROJECTION exploited species ENM/exploited_species_list.csv")

SAU_Data <- read.csv("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/SAU_Catch_Data_Test/SAU_Catch_Data_Test.csv")

#### Species per country

Spps <- SAU_Data %>% 
  filter(data_layer == "Reconstructed domestic catch") %>% 
  group_by(area_name,TaxonName = scientific_name) %>% 
  summarise(n()) %>% 
  left_join(exploited_species_list,
            by = "TaxonName")

# TakonKey <- Spps$TaxonKey[1:2]
TakonKey <- c("600004","600006",600107)
# Spp = 600988
TaxonKey <- unique(Spps$TaxonKey)

@@

NOTA: Explorando resultados y hay que arrelgar la cuestion de los tres modelos

```{Routine, eval = T, echo = F}

# Step 1; EstNeighbordList 
Neighbor_List <- EstNeighbors(Test)
# head(Neighbor_List)

# Step 2; GetSppDist
# Spp_Dist <- GetSppDist(TakonKey[1],"All")
# head(Spp_Dist)

# Multiple TaxonKeys
Spp_Dist <- NULL
Spp_Dist <- lapply(TaxonKey, FUN = GetSppDist, "All")

Spp_Dist <- Spp_Dist[!sapply(Spp_Dist, is.null)]
# Step 3; EstTransIndex
# Spp_Trans <- EstTransIndex(Spp_Dist,EEZ_Size,Neighbor_List)
# Spp_Trans

# Multiple TaxonKeys
Spp_Trans <- NULL
Spp_Trans <- bind_rows(lapply(Spp_Dist, FUN=EstTransIndex, EEZ_Size = EEZ_Size, Neighbors = Neighbor_List))

Spp_Trans %>% 
  mutate(Uncert = ifelse(Trans_Index >= 0.10 & Trans_Index <= 0.25,"10 %",
                         ifelse(Trans_Index >= 0.25 & Trans_Index <= 0.50,"25 %",
                                ifelse(Trans_Index >= 0.50,">50 %","No Trans")))
         ) %>% 
  group_by(Name,Uncert) %>% 
    summarise(nspp=n()) %>% 
  ggplot(.,
         aes(
           x = Name,
           y = nspp
         )
         )+
  geom_bar(stat ="identity") +
  facet_wrap(~Uncert)

@@

```{r World_Trans_Spp, eval=T,echo=T,warning=F,message=F}

@@

```{r Final_Analysis, eval=T, echo=F}

>>>>

```{r setup, eval = T, echo=F, warning=F,message=F, results='hide'}

@@

#### Library ####
packages <- c(
  "readxl", # Read dataframe
  "data.table", # Read dataframe (Fast!)
  "dplyr", # Data manipulation
  "tidyr", # Data manipulation
  "ggplot2", #Nice grpahs and spatial analysis
  # "cowplot",
  # "rgdal",
  # "RColorBrewer",
  # "knitr",
  # "kableExtra",
  # "ggrepel",
  # "gridExtra",
  # "ggmap",
  # "rgeos",
  # "stringr",
  # "stringer",
  "spdep", # for poly2nb
  "sf", #Spatial analysis 
  "sp", #Spatial analysis 
  "purrr",#Spatial analysis
  "rgdal", #Spatial analysis
  "tools", #Spatial analysis 
  # "png", # For reading plots in chunk codes
  # "grid" # For reading plots in chunk codes
  "parallel" # for parallelization
)

@@

# SAU <- length(list.files("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/SAU_Distribution/"))
# ENM <- length(list.files("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/ENM/"))
# Occ <- length(list.files("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/Occurence/"))

@@

#### List of species we have data for...

# Removes everything before the number
Step_One <- gsub(".*_","",list.files("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/SAU_Distribution/"))

# Species List
Species_List <- gsub("\\..*","",Step_One)

@@

EEZ_Size <- EEZ_CellID %>% 
  group_by(EEZID) %>% 
  summarise(n_tot_grid = n()) %>% 
  left_join(EEZIDs_List,
            by = "EEZID")

@@

  # SAU Distributions
  if(Model == "SAU_D"){
    File_Name <- paste("SAU_Distribution/DIST_GOOD_SP_")
  }

  # SAU Catch
  if(Model == "SAU_C"){
    File_Name <- paste("SAU_data_per_species/CATCH_SP_")
  }

@@

    Load <- lapply(Models_List, FUN=R.matlab::readMat, na.strings=0)

    sppdist <- as.data.frame(bind_cols(Load)) %>% 
      mutate(
        INDEX = INDEX,
        TaxonKey = Spp
      )
    colnames(sppdist) <- c("SAU_D","Occ","ENM","INDEX","TaxonKey")

    #### Step for SAU catch data that has to be averaged
    File_Name <- paste("SAU_data_per_species/CATCH_SP_")
    SppPath <- paste(Distpath,File_Name,Spp,".mat",sep="")

    # For now we're using only the last 10 years average, basicaaly if the species has been fished in any of these years, its considered present
    SAU_C_data <- as.data.frame(R.matlab::readMat(SppPath)) %>% 
      select(CATCH.56:CATCH.65) %>% 
      mutate(INDEX = INDEX) %>% 
      gather("Year","Catch",1:10) %>% # Last 10 years of data
      group_by(INDEX) %>% 
      summarise(SAU_C = mean(Catch,na.rm=T))

    # Join both tables
    sppdist <- sppdist %>% 
      left_join(SAU_C_data,
                by = "INDEX") %>% 
      select(TaxonKey,INDEX,everything())

@@

# Variables outside function
# head(GetSppDist(600004,"SAU_D")) # Works
# head(GetSppDist(600004,"SAU_C")) # Works
# head(GetSppDist(600004,"Occurence")) # Works
# head(GetSppDist(600004,"ENM")) # Works
# head(GetSppDist(600004,"All")) # Works

@@

```{r Trans_Spp_Function, eval= T ,echo = T,warning = F,message = F}

@@

# The function
EstTransIndex <- function(Spp,Model,EEZ_Size,Neighbors){

  # Get model data from spps
  SppDist <- GetSppDist(600004,"All")

@@

  Trans_Spp <- suppressWarnings(SppDist %>%
    gather("Model","Value",3:6) %>%
    filter(
      INDEX %in% Index_EEZ$INDEX,
      Value != 0
      ) %>%
    left_join(Neighbors,
              by = "INDEX") %>% # Match the species distribution to the Countries EEZs
    group_by(Name = Country_Territory,
             Neighbor_Territory,
             TaxonKey,
             Model
             ) %>% 
    summarise(n_cells_spp = n()) %>% # The number of cells present within each country's EEZ
    left_join(EEZ_Size,
              by = "Name") %>% 
    mutate(Trans_Index = n_cells_spp/n_tot_grid) %>%  # Include one transboundary species
    select(EEZID,everything())
    ) # For the sake of my brain

@@

### Testing Function ####
# Spp <- 600006
# Model <- "All"
# EEZ_Size <- EEZ_Size
# Neighbors <- Neighbor_List

@@

## Routine

@@

# Removes everything before the number
Step_One <- gsub(".*_","",list.files("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/SAU_Distribution/"))

# Species List
Species_List <- gsub("\\..*","",Step_One)

# Set Spceies list
Exploited_Species <- fread("/Volumes/DATA/PROJECTION exploited species ENM/exploited_species_list.csv") %>% 
  filter(TaxonKey %in% Species_List)

# For Function
Exploited_Species_List <- Exploited_Species$TaxonKey

####_________________________________________ TESTING PERIOD_________________________________________####
# I'm substetting central america to do the whole analysis (smaller EEZs = faster computation) once I have figure out things I'll run for all

@@

## Species Test Data Just to know what's captured in these countries 
# 
# SAU_Data <- fread("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/SAU_Catch_Data_Test/SAU_Catch_Data_Test.csv")
# # 
# # #### Species per country
# # 
# Spps <- SAU_Data %>%
#   filter(data_layer == "Reconstructed domestic catch") %>%
#   group_by(area_name,TaxonName = scientific_name) %>%
#   summarise(n()) %>%
#   left_join(Exploited_Species,
#             by = "TaxonName") %>% 
#   filter(TaxonKey %in% Exploited_Species_List$TaxonKey)
# 
# Exploited_Species_List <- Spps$TaxonKey

# Exploited_Species <- c("600004","600006",600107)
# Exploited_Species = 600004
# TaxonKey <- unique(Spps$TaxonKey)

@@

```{r Routine, eval = F, echo = F}

# Step 1; EstNeighbordList 
Neighbor_List <- EstNeighbors(Test)
# Neighbor_List <- EstNeighbors(World_EEZs)
# head(Neighbor_List)

# Step 2; EstTransIndex 
system.time(
  Spp_Trans <- bind_rows(
    lapply(
      Exploited_Species_List, FUN=EstTransIndex, EEZ_Size = EEZ_Size, Model = "All", Neighbors = Neighbor_List
    )
  )
)

@@

```{r Routine_parallel, eval = T, echo = F}

# Step 1; EstNeighbordList 
Neighbor_List <- EstNeighbors(Test)
# Neighbor_List <- EstNeighbors(World_EEZs)
# head(Neighbor_List)

# Check number of cores
n_Cores <- detectCores()
n_Cores

Spp_Trans <- NULL

system.time(
  Spp_Trans <- bind_rows(
    mclapply(
      Exploited_Species_List, FUN=EstTransIndex, EEZ_Size = EEZ_Size, Model = "All", Neighbors = Neighbor_List, mc.cores = n_Cores
    )
  )
)

### Save Test

# write.csv(Spp_Trans,
#           "central_america_trans.csv",
#           row.names = F)
# 
```

# Results

```{r, Trans_Results, eval = F, echo = F}

# Result 1. Number of tran spp per EEZ

Spp_Trans %>% 
  group_by(Name,TaxonKey) %>% 
  summarise(Mean_Index = mean(Trans_Index)) %>%  # The higher the less uncertain
  mutate(Uncert = ifelse(Mean_Index >= 0.50 ,"50 %",
                         ifelse(Mean_Index >= 0.25 ,"25 %",
                                ifelse(Mean_Index >= 0.10,">10 %","No Trans")))
         ) %>% 
  group_by(Name,Uncert) %>%
    summarise(nspp=n()) %>%
  ggplot(.,
         aes(
           x = Name,
           y = nspp#,
           # fill = Model
         )
         )+
  geom_bar(stat ="identity") +
  facet_wrap(~Uncert)

Spp_Trans %>% 
  group_by(Name,TaxonKey,Neighbor_Territory) %>% 
  summarise(Mean_Index = mean(Trans_Index)) %>%  # The higher the less uncertain
  mutate(Uncert = ifelse(Mean_Index >= 0.10 & Mean_Index <= 0.25,"10 %",
                         ifelse(Mean_Index >= 0.25 & Mean_Index <= 0.50,"25 %",
                                ifelse(Mean_Index >= 0.50,">50 %","No Trans")))
         ) %>% 
    group_by(Name,Neighbor_Territory,Uncert) %>% 
    summarise(nspp=n())

```{r World_Trans_Spp, eval = F,echo = T,warning = F,message = F}

@@

```{r Final_Analysis, eval = F, echo=F}
